---
title: "R-course"
format: html
editor: 
  markdown: 
    wrap: 72
---

What did we do in the R course?

First we downloaded RStudio and got to know the interface with its different panels. We did basic calculations, defined variables and used different mathematical functions. 
The first exercise looked like this:

```{r}

x1 <- sqrt(1^2+2^2+3^2+4^2)

list(x1)
rm(x1)
```

Next we learned about logical operators and how to use them. We also dealt with different data structures, vectors and vector functions:

```{r}
x <- seq(from=10, to=20, by=0.5)
length(x)
y <- seq(from=100, to=200, length=21)
mean(x)
mean(y)
z <- data.frame(x,y)
z
x[8]
y[8]
```

Following that we looked at matrices and matrices functions:

```{r}
h <- seq(1, 4, by=0.2)
length(h)
H <- matrix(h, nrow=4, byrow=TRUE)
colMeans(H)
```

## Example for linear regression

```{r}
set.seed(123)

n <- 100 
x <- runif(n, 0, 10)
y <- 5 + 2 * x + rnorm(n, 0, 2)
```

Fit linear model

```{r}
fit <- lm(y ~ x)
summary(fit)
```

Extract parameters

```{r}
beta_hat <- coef(fit) 
sigma_hat <- summary(fit)$sigma
```

Bootstrap

```{r}
B <- 1000
beta_boot <- matrix(NA, nrow = B, ncol = length(beta_hat))

for (b in 1:B) {
  y_star <- beta_hat[1] + beta_hat[2] * x + rnorm(n, 0, sigma_hat)
  fit_star <- lm(y_star ~ x)
  beta_boot[b, ] <- coef(fit_star)
}
```

Results

```{r}
boot_means <- colMeans(beta_boot)
boot_ses <- apply(beta_boot, 2, sd)
boot_ci <- apply(beta_boot, 2, quantile, probs = c(0.025, 0.975))

boot_means
boot_ses
boot_ci
```

## Example for logistic regression

Generate data using runif (randomize uniform distribution)

```{r}
set.seed(123)

n <- 200
x <- runif(n, 0, 10)

beta0 <- -2
beta1 <- 0.5

p <- 1 / (1 + exp(-(beta0 + beta1 * x)))
y <- rbinom(n, 1, p)

data <- data.frame(y = y, x = x)

fit <- glm(y ~ x,
           family = binomial(link = "logit"),
           data = data)

summary(fit)
```

Extract estimated parameters

```{r}
beta_hat <- coef(fit)

B <- 1000
beta_boot <- matrix(NA, nrow = B, ncol = length(beta_hat))


for (b in 1:B) {
  p_hat <- 1 / (1 + exp(-(beta_hat[1] + beta_hat[2] * x)))
  y_star <- rbinom(n, 1, p_hat)
  
  fit_star <- glm(y_star ~ x,
                  family = binomial(link = "logit"))
  
  beta_boot[b, ] <- coef(fit_star)
}

colnames(beta_boot) <- names(beta_hat)
```

Results

```{r}
boot_means <- colMeans(beta_boot)
boot_ses <- apply(beta_boot, 2, sd)
boot_ci <- apply(beta_boot, 2, quantile, probs = c(0.025, 0.975))

results <- data.frame(
  Estimate = beta_hat,
  BootMean = boot_means,
  BootSE = boot_ses,
  CI_lower = boot_ci[1, ],
  CI_upper = boot_ci[2, ]
)

print(results)
```

lets graph both intercept and slope using histogram

```{r}
par(mfrow = c(1, 2))

hist(beta_boot[,1],
     main = "Bootstrap Intercept",
     xlab = "beta0*",
     breaks = 30)
abline(v = beta_hat[1], col = "red", lwd = 2)

hist(beta_boot[,2],
     main = "Bootstrap Slope",
     xlab = "beta1*",
     breaks = 30)
abline(v = beta_hat[2], col = "red", lwd = 2)
```
